---
title: "Mini-Project 2"
author: "Sarah Wu, Junru Wu, Irene Xu"
date: "2019-03-24"
output: 
  html_document:
    code_folding: hide
---
```{r include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggthemes)
```
### Introduction 
* The present project is based on the datasets provided by the [Federal Election Committee](https://classic.fec.gov/finance/disclosure/ftpdet.shtml#archive_link) that include data about the various candidates, committees, contributions from committees to candidates they support, and House Elections results in 2012. 
* We were interested in how candidates’ party affiliation and the type of entity making the transactions affect the relationship between the amount of transaction spent (couldn't find the unit, so we assumed in dollars) to support the candidates and the general votes they received.
* We hypothesized the relationship would be affected due to the complex interactions among differences in the philosophy, political and economic strategies that different parties and different entities have, as well as the variety in their scope of influence.

#### 1. Loading data
```{r, message = FALSE, warning = FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```
#### 2. Data wrangling
+ We kept the house candidates affiliated with the Democratic Party or the Republican Party (because most audience are interested in them) and kept the variables indicating their id and party affiliation in a way that's easier to be understood.
```{r message = FALSE, warning = FALSE}
candidates_2 <- candidates %>%
  filter(cand_party_affiliation %in% c("DEM", "REP")) %>%
  filter(cand_office == "H") %>%
  mutate(cand_party = ifelse(cand_party_affiliation == "DEM", "Democratic Party", "Republican Party")) %>%
  select(cand_id, cand_party) 
```
+ In the contributions dataset, we were interested in the amount of transaction committees used to support candidates.
```{r message = FALSE, warning = FALSE}
contributions_2 <- contributions %>%
  filter(transaction_amt > 0) %>%
  select(cand_id, transaction_amt)
```
+ In the House Election in 2012, we chose the candidates who were affiliated with the Democrate Party or the Repulican Party and did receive votes.
```{r message = FALSE, warning = FALSE}
house_2012 <- house_elections %>%
  filter(party %in% c("D", "R")) %>%
  select(fec_id, general_votes, -party) %>% 
  filter(general_votes > 0)
```
+ We joined the three processed tables by candidates' id and calculated the total amount of transaction they received. 
```{r message = FALSE, warnings = FALSE}
join_1 <- candidates_2 %>%
  inner_join(house_2012, by = c("cand_id" = "fec_id")) %>%
  inner_join(contributions_2, by = c("cand_id" = "cand_id")) %>%
  group_by(cand_id) %>%
  mutate(total_tran = sum(transaction_amt))
join_2 <- distinct(.data = join_1, cand_id, .keep_all = TRUE)
```
#### 3. Ploting the relationship between the amount of contribution used to support House candidates and the general votes they received, separated by their party affiliation. 
+ In the first graph, we kept all the data, with the dotted lines indicating the means of relative measures and the blue regression line.
```{r message = FALSE, warning = FALSE}
ggplot(join_2, aes(x = total_tran, y = general_votes)) +
  geom_point(alpha = 0.16, color = "red") +
  ggtitle("Is the money spent on House candidates really worth it?") +
  scale_x_continuous("Total amount spent in suppoting") +
  scale_y_continuous("General Votes") +
  facet_wrap(~ cand_party) +
  geom_smooth(method = lm, se = FALSE, color = "dodgerblue2", size = 1) +
  geom_line(y = mean(join_2$general_votes), linetype = 3, size = 0.75) +
  geom_vline(xintercept = mean(join_2$total_tran), linetype = 3, size = 0.75) +
  theme_minimal()
```

+ Since there are some data far on the x-axis that don't have much representative or predictive value, we put limits on the second graph to reduce their effects.
```{r message = FALSE, warning = FALSE}
ggplot(join_2, aes(x = total_tran, y = general_votes)) +
  geom_point(alpha = 0.16, color = "red") +
  ggtitle("Is the money spent on House candidates really worth it?") +
  scale_x_continuous("Total amount spent to support a candidate", limits = c(0, 2000000), labels = c("0", "0.5M", "1M", "1.5M", "2M")) +
  scale_y_continuous("General votes a candidate received", limits = c(0, 300000), labels = c("0", "100k", "200k", "300k")) +
  facet_wrap(~ cand_party) +
  geom_smooth(method = lm, se = FALSE, color = "dodgerblue2", size = 1) +
  geom_line(y = mean(join_2$general_votes), linetype = 3, size = 0.75) +
  geom_vline(xintercept = mean(join_2$total_tran), linetype = 3, size = 0.75) +
  theme_minimal()
```

##### Analysis
+ According to both graphs, there is a positive relationship between the amount of transaction spent to support a candidate and the general votes they received. However, the party affiliation of candidates doesn't have much influence on this relationship.

#### 4. Next, we wanted to see whether the type of entity making the transaction has an effect on this relationship. 
+ We created a function that calculates the total amount of transaction for each candidate made by a particular type of entity.
```{r message = FALSE, warning = FALSE}
entity_type <- function(e_type) {
contributions %>%
    filter(entity_type == e_type) %>%
    group_by(cand_id) %>%
    summarise(transaction_total = sum(transaction_amt, na.rm = TRUE))   %>%
    inner_join(house_elections, by = c("cand_id" = "fec_id"))   %>%
    select(cand_id, transaction_total, general_votes)
}
```
+ Another function that plots the relationship between the amount a type of entity spent in supporting candidates and the general votes these candidates received in the 2012 House Election. (With limits as twice of the means on both axes for clearer visulation and less influence by potential outliers/leverage points)
```{r message = FALSE, warning = FALSE}
plot_by_type <- function(e_type){
ggplot(entity_type(e_type), aes(x = transaction_total, y = general_votes)) +
  geom_point(alpha = 0.16, color = "red") +
  ggtitle(e_type) +
  scale_x_continuous("Total amount spent in suppoting", limits = c(0, 2 * mean(entity_type(e_type)$transaction_total))) +
  scale_y_continuous("General Votes", limits = c(0, 2 * mean(entity_type(e_type)$general_votes))) +
  geom_smooth(method = lm, se = FALSE, color = "dodgerblue2", size = 1) +
  geom_line(y = mean(entity_type(e_type)$general_votes), linetype = 3, size = 0.75) +
  geom_vline(xintercept = mean(entity_type(e_type)$transaction_total), linetype = 3, size = 0.75) +
  theme_minimal()
}
```

#### 5. We plotted the relationship for all 7 types of eneities. 
####     Candidate (CAN)
```{r message = FALSE, warning = FALSE}
plot_by_type("CAN")
```

####     Candidate Committee (CCM)
```{r message = FALSE, warning = FALSE}
plot_by_type("CCM")
```

####     Committee (COM)
```{r message = FALSE, warning = FALSE}
plot_by_type("COM")
```

####     Individual (IND)
```{r message = FALSE, warning = FALSE}
plot_by_type("IND")
```

####     Political Action Committee (PAC)
```{r message = FALSE, warning = FALSE}
plot_by_type("PAC")
```

####     Organization (ORG) (not a committee and not a person)
```{r message = FALSE, warning = FALSE}
plot_by_type("ORG")
```

####     Party Organization (PTY)
```{r message = FALSE, warning = FALSE}
plot_by_type("PTY")
```

#### Analysis
* From the graphs we generated above, we found different relationships between the general votes candidates received and the transaction amount spent by different types of entity to support the House candidates. 
* Though the data are very scattered, we did find the number of general votes candidates received has:
* (1) a moderately positive relationship with the transactions made by Candidates (CAN), Candidate Committees (CCM) and Committees (COM), 
* (2) weakly positive relationships with those made by Political Action Committees (PAC) and Individuals (IND), and 
* (3) weakly negative relationships with those made by Organizations (ORG) and Party Organizations (PTY). 
* Therefore, the transactions made by Candidates (CAN), Candidates Committees (CCM), and Committees (COM) could be considered more effective than those made by the other 4 types of entity if we can rule out possible confounding variables.

#### General Conclusion
*  From out previous data, graphs and analysis, we can conclude that in the 2012 U.S House Election, the positive relationship between transaction amounts used to support candidates and the general votes they received did not vary much by candidates’ party affiliations (only for Democratic and Republican party in this case), but was influenced by the type of entity that made the transactions.

> Word count: `r wordcountaddin::word_count()`

^[[Github Repository](https://github.com/jwu13/sds192-mp2)]

^[* The functions we created can be applied to other elections (i.e. the 2015 U.S Presidential Election) that has different types of entity making transactions, to other entity types, and to candidates with other party affiliations.]

^[* Data retrieved from https://classic.fec.gov/finance/disclosure/ftpdet.shtml#archive_link]
